A quantitative Mac Diarmid's inequality for geometrically ergodic Markov chains

### 2019-12-30 Note_exponential_moment/MarkovCon.tex
update

For any $\cB\in \cX$ and any integer $i\geqslant 0$, let
\[
\tau_{\cB}^i=\inf\{n\geqslant i: X_n\in \cB\}=i+\tau_{\cB}^0\circ \theta^i
\qquad \text{and} \qquad
\sigma_{\cB}=\tau_{\cB}^1=1+\tau_{\cB}^0\circ \theta\eqsp.
\]
For $c = \chunk[0]{c}[n-1] \in \R_+^n$, we denote by  $\functionboundeddiff{\sX^n}{c}$  the set of measurable functions $f:\sX^n\to\R$ such that for all $x= (x_0,\dots,x_{n-1})$ and $y= (y_0,\dots,y_{n-1})$, $|f(x)-f(y)|\leqslant \sum_{i=0}^{n-1} c_i\indiacc{x_i\ne y_i}$
The main result is established under the following conditions.
\begin{itemize}
\item[{\bf H1}] The Markov kernel $P$ is irreducible and aperiodic, with unique invariant probability $\pi$.
\item[{\bf H2}] There exist a non-empty set $\cC \in \cX$ and two real numbers $u>1$ and $M>0$ such that
  \[\sup_{x\in \cC}\Ebb_x[u^{\sigma_{\cC}}]\leqslant M\eqsp.\]
\item[{\bf H3}] There exist $r\in (0,1)$ and $L\geqslant 1$ such that, for any $x$ in the set $\cC$ of {\bf H2} and any $n\geqslant 0$,
\[
\tvdist{\delta_xP^n}{\pi} \leqslant Lr^n\eqsp,
\]
where $\pi$ is the unique invariant measure granted in {\bf H1}.
\end{itemize}
When the Markov kernel $P$ is uniformly ergodic, then {\bf H3} holds with $\cC=\sX$. The following Lemma is a coupling result that replaces \cite[Lemma~6]{MR3407208}.
It is instrumental in the sequel.
\begin{lemma}\label{lem:DMPS18}
 For any probability measures $\xi$ and $\xi'$ on $(\sX,\cX)$, any $n\geqslant 1$, any $c\in \R_+^n$ and any $h\in \bB\Dbb(\sX^n,c)$,
\begin{equation*}\label{eq:Step1}
 |\Ebb_\xi[h(\chunk[0]{X}[n-1])]-\Ebb_{\xi'}[h(\chunk[0]{X}[n-1])]|\leqslant 2\sum_{i=0}^{n-1}c_i \tvdist{\xi P^i}{\xi'P^i}\eqsp.
\end{equation*}
\end{lemma}
\begin{remark}
 It is possible to avoid the factor $2$ in \eqref{eq:Step1} under additional technical conditions, for example, when there exists a maximal coupling for $(\Pbb_\xi,\Pbb_{\xi'})$, see \cite[Lemma 23.2.1]{douc:moulines:priouret:soulier:2018}.
\end{remark}
\begin{proof}
Fix an arbitrary
$x^* \in \Xset$. For $i \in \{1,\ldots,n-1\}$, we set $\bar{h}_i(\chunk[i]{x}[n-1])=h(x^*,\dots,x^*,\chunk[i]{x}[n-1])$. By convention,
we set $\bar{h}_n$ the constant function $\bar{h}_n=h(x^*,\ldots,x^*)$ and $\bar{h}_0=h$. With these
notations, we have the decomposition
\[
h(\chunk{x})=\sum_{i=0}^{n-1} \{\bar{h}_i(\chunk[i]{x}[n-1]) - \bar{h}_{i+1}(\chunk[i+1]{x}[n-1]) \} +\bar{h}_n \eqsp.
\]
For all $i \in \{0,\ldots,n-1\}$ and all $x_i \in \Xset$, let
\begin{align}
\bar{w}_i(x_i)&=\int \lrcb{\bar{h}_i(\chunk[i]{x}[n-1])-\bar{h}_{i+1}(\chunk[i+1]{x}[n-1])}  \prod_{\ell=i+1}^{n-1} P(x_{\ell-1}, \rmd x_{\ell})\eqsp, \nonumber \\
&=\int \lrcb{h(x^*,\ldots,x^*,\chunk[i]{x})-h(x^*,\ldots,x^*,\chunk[i+1]{x})}  \prod_{\ell=i+1}^{n-1} P(x_{\ell-1}, \rmd x_{\ell}) \eqsp. \label{eq:delta:diarmid:fond}
\end{align}
It is easily seen that $\CPE{\{\bar{h}_i(\chunk[i]{X}[n-1])-\bar{h}_{i+1}(\chunk[i+1]{X}[n-1])\}}{\cF_{i}}=\bar{w}_i(X_i)$, $\Pbb_\xi-\as$,  which implies that
\[
\Ebb_\xi \lrb{h(\chunk{X})}=\sum_{i=0}^{n-1} \xi P^i \bar{w}_i + \bar{h}_n \eqsp.
\]
Since  $h \in\functionboundeddiff{\Xset^n}{c}$, \eqref{eq:delta:diarmid:fond} shows that  $\supnorm{\bar{w}_i} \leq c_i$.
Therefore,
\[
|\Ebb_\xi \lrb{h(X^{n-1})}-\Ebb_{\xi'} \lrb{h(X^{n-1})} |\\
\leq \sum_{i=0}^{n-1} |\xi P^i \bar{w}_i-\xi' P^i \bar{w}_i| \leq 2 \sum_{i=0}^{n-1} c_{i}\tvdist{\xi P^{i}}{\xi' P^{i}} \eqsp.
\]
\end{proof}

\section{Main result}\label{sec:Main}
The main result of this paper  is the following quantitative version of Mac Diarmid's inequality for geometrically ergodic Markov chains.
\begin{theorem}\label{thm:ConcMarkQuant}
 Assume {\bf H1}, {\bf H2}, {\bf H3}. Let $n\geqslant 1$, $c\in \R^n$ and  $f\in \functionboundeddiff{\sX^n}{c}$. Then, for all $x\in \cC$ and $t > 0$,
 \[
\Pbb_x\big(f( \chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]>t\big)\leqslant \exp\bigg(-\frac{\beta t^2}{\|c\|^2}\bigg)\eqsp,
 \]
where $\beta$ is given by
 \[
 \beta=\frac{(1-r\vee u^{-1/4})^2}{16L}\bigg(\frac5{\log u}+4ML\bigg)^{-1}\eqsp.
 \]
\end{theorem}
\begin{proof}[Proof of \Cref{thm:ConcMarkQuant}]
Fix $c\in \Rbb^n$, $x\in\sX$ and $f\in \functionboundeddiff{\sX^n}{c}$.
Following \cite{MR3407208}, we decompose $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$ into martingale increments by conditioning to the stopping times $\tau_{\cC}^i$, $i=0,\ldots, n-1$.
For any integer $i\in [0,n-1]$, define
\[
G_i=\Ebb_x\big[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^i}\big]\eqsp.
\]
As $\tau_{\cC}^0=0$ $\Pbb_x$-a.s., it holds $\Ebb_x[f(\chunk[0]{X}[n-1])]=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^0}]=G_0$.
Moreover, as $\tau_{\cC}^{n-1}\geqslant n-1$, it also holds $G_{n-1}=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{n-1}}]=f(\chunk[0]{X}[n-1])$.
Therefore, the difference $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$ is decomposed into a sum of the martingale increments $G_{i+1}-G_i$ as follows
\begin{equation}\label{eq:LinkWithG}
 f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]=G_{n-1}-G_0=\sum_{i=0}^{n-2}(G_{i+1}-G_i)\eqsp.
\end{equation}
The proof is now decomposed into three facts that aim at bounding the Laplace transform of $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$.

@@

\begin{lemma}
\label{lem:gi-gipi}
 Assume {\bf H1}, {\bf H2}, {\bf H3}. 
 For any $i\in \{0,\ldots,n-1\}$ and $(\chunk[0]{x}[i-1],x_i)$ in $\sX^i\times\cC$,
\begin{equation}\label{eq:gi-gipi}
|g_i(\chunk[0]{x}[i])-g_{i,\pi}(\chunk[0]{x}[i])|\leqslant 2L\sum_{j=i+1}^{n-1}c_jr^{j-i}\eqsp.
\end{equation}
\end{lemma}
\begin{proof}
 Fix $i\in\{0,\ldots,n-1\}$ and $\chunk[0]{x}[i]\in \sX^{i+1}$.
As $f\in\functionboundeddiff{\sX^n}{c}$, the function $\tilde{f}_i:\chunk[1]{y}[n-1-i]\in \sX^{n-1-i}\mapsto f(\chunk[0]{x}[i],\chunk[1]{y}[n-1-i])\in \R$ satisfies
\[
|\tilde{f}_i(\chunk[1]{y}[n-1-i])-\tilde{f}_i(\chunk[1]{z}[n-1-i])|\leqslant \sum_{k=1}^{n-1-i}c_{i+k}\indiacc{y_k\ne z_k}\eqsp.
\]
Hence, $\tilde{f}_i\in\bB\Dbb(\sX^{n-1-i},c_{i+1:n-1})$. Applying Lemma~\ref{lem:DMPS18} to the function $h=\tilde{f}_i$ yields
\begin{align*}
 |g_i(\chunk[0]{x}[i])-g_{i,\pi}(\chunk[0]{x}[i])|&=|\Ebb_{x_i}[f(\chunk[0]{x}[i],\chunk[1]{X}[n-1-i])]-\Ebb_{\pi}[f(\chunk[0]{x}[i],\chunk[1]{X}[n-1-i])]|\\
 &=|\Ebb_{x_i}[\tilde{f}_i(\chunk[1]{X}[n-1-i])]-\Ebb_{\pi}[\tilde{f}_i(\chunk[1]{X}[n-1-i])]| \leqslant 2\sum_{j=i+1}^{n-1}c_j \tvdist{\delta_{x_i}P^j}{\pi}\eqsp.
\end{align*}
Inequality \eqref{eq:gi-gipi} follows from {\bf H3}.
\end{proof}

>>>>

For any $\cB\in \cX$ and any integer $i\geqslant 0$, let
\tau_{\cB}^i=\inf\{n\geqslant i: X_n\in \cB\}=i+\tau_{\cB}^0\circ \theta^i
\qquad \text{and} \qquad
\sigma_{\cB}=\tau_{\cB}^1=1+\tau_{\cB}^0\circ \theta\eqsp.
For $c = \chunk[0]{c}[n-1] \in \R_+^n$, we denote by  $\functionboundeddiff{\sX^n}{c}$  the set of measurable functions $f:\sX^n\to\R$ such that for all $x= (x_0,\dots,x_{n-1})$ and $y= (y_0,\dots,y_{n-1})$, $|f(x)-f(y)|\leqslant \sum_{i=0}^{n-1} c_i\indiacc{x_i\ne y_i}$
The main result is established under the following conditions.
\begin{itemize}
\item[{\bf H1}] The Markov kernel $P$ is irreducible and aperiodic, with unique invariant probability $\pi$.
\item[{\bf H2 (drift condition)}] There exist a measurable function $V: X \rightarrow [1; +\infty)$, real numbers $\lambda \in (0,1)$, $b > 0$ and a non-empty set $\cC \in \cX$ such that for any $x \in X$,
PV(x) \leq \lambda V(x) + b\indiacc{x \in \cC}
%\item[{\bf H2}] There exist a non-empty set $\cC \in \cX$ and two real numbers $u>1$ and $M>0$ such that
%  \[\sup_{x\in \cC}\Ebb_x[u^{\sigma_{\cC}}]\leqslant M\eqsp.\]
%\item[{\bf H3}] There exist $r\in (0,1)$ and $L\geqslant 1$ such that, for any $x$ in the set $\cC$ of {\bf H2} and any $n\geqslant 0$,
%\[
%\tvdist{\delta_xP^n}{\pi} \leqslant Lr^n\eqsp,
%\]
%where $\pi$ is the unique invariant measure granted in {\bf H1}.
\item[{\bf H3 (V-ergodicity)}] There exist $r \in (0,1)$, $L \geqslant 1$ and function $V: X \rightarrow \mathbb{R}$ such that for any $x  \in  X$ and any $n \geqslant 0$, it holds
d_V(\delta_xP^n,\pi) \leqslant L V(x) r^n\eqsp,
\end{itemize}
where $\pi$ is the unique invariant measure from {\bf H1}.
\par
When the Markov kernel $P$ is uniformly ergodic, then {\bf H3} holds with $V(x) = 1$.   
\par
Note that \cite[Proposition 14.3.1]{douc:moulines:priouret:soulier:2018} implies that if Markov kernel $P$ satisfies the drift condition {\bf H2} with the set $\cC \in \cX$, $\cC = \{x: V(x) < d\}$, then it holds for any $x \in \cC$ that
\sup_{x\in \cC}\Ebb_x[\lambda^{-\sigma_{\cC}}]\leqslant M\eqsp
with some $M$ depending on $V(x)$ and $d$.
\par
The following Lemma is a coupling result that replaces \cite[Lemma~6]{MR3407208}. It is instrumental in the sequel.
\begin{lemma}\label{lem:DMPS18}
 For any probability measures $\xi$ and $\xi'$ on $(\sX,\cX)$, any $n\geqslant 1$, any $c\in \R_+^n$ and any $h\in \bB\Dbb(\sX^n,c)$,
\begin{equation*}\label{eq:Step1}
 |\Ebb_\xi[h(\chunk[0]{X}[n-1])]-\Ebb_{\xi'}[h(\chunk[0]{X}[n-1])]|\leqslant 2\sum_{i=0}^{n-1}c_i \tvdist{\xi P^i}{\xi'P^i} \leqslant 2\sum_{i=0}^{n-1}c_i \vfuncdist{\xi P^i}{\xi'P^i} \eqsp.
\end{equation*}
\end{lemma}
\begin{remark}
 It is possible to avoid the factor $2$ in \eqref{eq:Step1} under additional technical conditions, for example, when there exists a maximal coupling for $(\Pbb_\xi,\Pbb_{\xi'})$, see \cite[Lemma 23.2.1]{douc:moulines:priouret:soulier:2018}.
\end{remark}
\begin{proof}
Fix an arbitrary
$x^* \in \Xset$. For $i \in \{1,\ldots,n-1\}$, we set $\bar{h}_i(\chunk[i]{x}[n-1])=h(x^*,\dots,x^*,\chunk[i]{x}[n-1])$. By convention,
we set $\bar{h}_n$ the constant function $\bar{h}_n=h(x^*,\ldots,x^*)$ and $\bar{h}_0=h$. With these
notations, we have the decomposition
h(\chunk{x})=\sum_{i=0}^{n-1} \{\bar{h}_i(\chunk[i]{x}[n-1]) - \bar{h}_{i+1}(\chunk[i+1]{x}[n-1]) \} +\bar{h}_n \eqsp.
For all $i \in \{0,\ldots,n-1\}$ and all $x_i \in \Xset$, let
\bar{w}_i(x_i)&=\int \lrcb{\bar{h}_i(\chunk[i]{x}[n-1])-\bar{h}_{i+1}(\chunk[i+1]{x}[n-1])}  \prod_{\ell=i+1}^{n-1} P(x_{\ell-1}, \rmd x_{\ell})\eqsp, \nonumber \\
&=\int \lrcb{h(x^*,\ldots,x^*,\chunk[i]{x})-h(x^*,\ldots,x^*,\chunk[i+1]{x})}  \prod_{\ell=i+1}^{n-1} P(x_{\ell-1}, \rmd x_{\ell}) \eqsp. \label{eq:delta:diarmid:fond}
It is easily seen that $\CPE{\{\bar{h}_i(\chunk[i]{X}[n-1])-\bar{h}_{i+1}(\chunk[i+1]{X}[n-1])\}}{\cF_{i}}=\bar{w}_i(X_i)$, $\Pbb_\xi-\as$,  which implies that
\Ebb_\xi \lrb{h(\chunk{X})}=\sum_{i=0}^{n-1} \xi P^i \bar{w}_i + \bar{h}_n \eqsp.
Since  $h \in\functionboundeddiff{\Xset^n}{c}$, \eqref{eq:delta:diarmid:fond} shows that  $\supnorm{\bar{w}_i} \leq c_i$.
|\Ebb_\xi \lrb{h(X^{n-1})}-\Ebb_{\xi'} \lrb{h(X^{n-1})} |\\
\leq \sum_{i=0}^{n-1} |\xi P^i \bar{w}_i-\xi' P^i \bar{w}_i| \leq 2 \sum_{i=0}^{n-1} c_{i}\tvdist{\xi P^{i}}{\xi' P^{i}} \eqsp.
The result now follows from the trivial bound $\tvdist{\xi}{\eta} \leq \vfuncdist{\xi}{\eta}$ for arbitrary measures $\xi,\eta$ given that $V(x) \geq 1$ for any $x \in X$.

\section{Main result}\label{sec:Main}
The main result of this paper  is the following quantitative version of Mac Diarmid's inequality for $V-$ergodic Markov chains under certain drift condition.
%\begin{theorem}\label{thm:ConcMarkQuant}
 %Assume {\bf H1}, {\bf H2}, {\bf H3}. Let $n\geqslant 1$, $c\in \R^n$ and  $f\in \functionboundeddiff{\sX^n}{c}$. Then, for all $x\in \cC$ and $t > 0$,
 %\[
%\Pbb_x\big(f( \chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]>t\big)\leqslant \exp\bigg(-\frac{\beta t^2}{\|c\|^2}\bigg)\eqsp,
% \]
%where $\beta$ is given by
 %\[
 %\beta=\frac{(1-r\vee u^{-1/4})^2}{16L}\bigg(\frac5{\log u}+4ML\bigg)^{-1}\eqsp.
 %\]
%\end{theorem}
\begin{theorem}\label{thm:ConcMarkVErgodic}
 Assume {\bf H1}, {\bf H2} and {\bf H3} holds with function $V(x)$. Let $\cC = \{V(x) < d\}$, $n\geqslant 1$, $c\in \R^n$ and  $f\in \functionboundeddiff{\sX^n}{c}$. Then, for all $x\in \cC$ and $t > 0$,
 \[
\Pbb_x\big(f( \chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]>t\big)\leqslant \exp\bigg(-\frac{\beta t^2}{\|c\|^2}\bigg)\eqsp,
 \]
where $\beta$ is given by
 \[
 \beta=\frac{(1-r\vee \lambda^{1/4})^2}{16Ld}\bigg(-\frac5{\log \lambda}+4MLd\bigg)^{-1}\eqsp.
 \]
\end{theorem}

\begin{proof}[Proof of \Cref{thm:ConcMarkVErgodic}]
Fix $c\in \Rbb^n$, $x\in\sX$ and $f\in \functionboundeddiff{\sX^n}{c}$. Note that drift condition {\bf H3} implies that
\[
\sup_{x\in \cC}\Ebb_x[\lambda^{-\sigma_{\cC}}]\leqslant M\eqsp
\]
Following \cite{MR3407208}, we decompose $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$ into martingale increments by conditioning to the stopping times $\tau_{\cC}^i$, $i=0,\ldots, n-1$.
For any integer $i\in [0,n-1]$, define
\[
G_i=\Ebb_x\big[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^i}\big]\eqsp.
\]
As $\tau_{\cC}^0=0$ $\Pbb_x$-a.s., it holds $\Ebb_x[f(\chunk[0]{X}[n-1])]=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^0}]=G_0$.
Moreover, as $\tau_{\cC}^{n-1}\geqslant n-1$, it also holds $G_{n-1}=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{n-1}}]=f(\chunk[0]{X}[n-1])$.
Therefore, the difference $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$ is decomposed into a sum of the martingale increments $G_{i+1}-G_i$ as follows
\begin{equation}\label{eq:LinkWithG}
 f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]=G_{n-1}-G_0=\sum_{i=0}^{n-2}(G_{i+1}-G_i)\eqsp.
\end{equation}
The proof is now decomposed into three facts that aim at bounding the Laplace transform of $f(\chunk[0]{X}[n-1])-\Ebb_x[f(\chunk[0]{X}[n-1])]$.

@@

\begin{lemma}
\label{lem:gi-gipi}
 Assume {\bf H1}, {\bf H2}, {\bf H3} and let $\cC = \{x:V(x) < d\}$. 
 For any $i\in \{0,\ldots,n-1\}$ and $(\chunk[0]{x}[i-1],x_i)$ in $\sX^i\times\cC$,
\begin{equation}\label{eq:gi-gipi}
|g_i(\chunk[0]{x}[i])-g_{i,\pi}(\chunk[0]{x}[i])|\leqslant 2LV(x_i)\sum_{j=i+1}^{n-1}c_jr^{j-i} \leqslant 2Ld \sum_{j=i+1}^{n-1}c_jr^{j-i}\eqsp.
\end{equation}
\end{lemma}
\begin{proof}
 Fix $i\in\{0,\ldots,n-1\}$ and $\chunk[0]{x}[i]\in \sX^{i+1}$.
As $f\in\functionboundeddiff{\sX^n}{c}$, the function $\tilde{f}_i:\chunk[1]{y}[n-1-i]\in \sX^{n-1-i}\mapsto f(\chunk[0]{x}[i],\chunk[1]{y}[n-1-i])\in \R$ satisfies
\[
|\tilde{f}_i(\chunk[1]{y}[n-1-i])-\tilde{f}_i(\chunk[1]{z}[n-1-i])|\leqslant \sum_{k=1}^{n-1-i}c_{i+k}\indiacc{y_k\ne z_k}\eqsp.
\]
Hence, $\tilde{f}_i\in\bB\Dbb(\sX^{n-1-i},c_{i+1:n-1})$. Applying Lemma~\ref{lem:DMPS18} to the function $h=\tilde{f}_i$ yields
\begin{align*}
 |g_i(\chunk[0]{x}[i])-g_{i,\pi}(\chunk[0]{x}[i])|&=|\Ebb_{x_i}[f(\chunk[0]{x}[i],\chunk[1]{X}[n-1-i])]-\Ebb_{\pi}[f(\chunk[0]{x}[i],\chunk[1]{X}[n-1-i])]|\\
 &=|\Ebb_{x_i}[\tilde{f}_i(\chunk[1]{X}[n-1-i])]-\Ebb_{\pi}[\tilde{f}_i(\chunk[1]{X}[n-1-i])]| \leqslant 2\sum_{j=i+1}^{n-1}c_j \vfuncdist{\delta_{x_i}P^{j-i}}{\pi} \eqsp.
\end{align*}
Now inequality \eqref{eq:gi-gipi} follows from {\bf H3}.
\end{proof}

\noindent
{\bf Fact 2.} \emph{Let $\rho$ such that $r\leqslant \rho<1$ and $i\in\{1,\ldots,n-1\}$.
Then,}
\begin{align}
\label{eq:Gi-Gi-1} |G_i-G_{i-1}|&\leqslant C_1\|c\|_{\infty}\indiacc{\tau_{\cC}^{i-1}=i-1}\sigma_{\cC}\circ\theta^{i-1}\eqsp,\\
\label{eq:Gi-Gi-1Sq}  |G_i-G_{i-1}|^2&\leqslant C_2\indiacc{\tau_{\cC}^{i-1}=i-1}\frac1{\rho^{2\sigma_{\cC}\circ\theta^{i-1}}}\sum_{k=i}^{n-1}c_k^2\rho^{k-i}\eqsp.
\end{align}
\emph{where, $C_1=5Ld/(1-r)$ and $C_2=16L^2d^2/(1-\rho)$.}
\begin{proof}[Proof of Fact 2.]
For any integer $i\in\{1,\ldots,n\}$, let
\[
G_{i,1}=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{i-1}}]\indiacc{\tau_{\cC}^{i-1}=i-1},\qquad G_{i,2}=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{i}}]\indiacc{\tau_{\cC}^{i-1}=i-1}\eqsp.
\]
From Fact 1., $G_i-G_{i-1}=G_{i,2}-G_{i,1}$.
By Markov's property, for any $i\in \{0,\ldots,n-1\}$ and  $x\in \sX$,
\[
\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_i]=g_i(X_{0:i}),\qquad \Pbb_{x}-\text{a.s.}\eqsp.
\]
Now, let $R_{i,1}=g_{i-1}(\chunk[0]{X}[i-1])\indiacc{\tau_{\cC}^{i-1}=i-1}-g_{i-1,\pi}(\chunk[0]{X}[i-1])\indiacc{\tau_{\cC}^{i-1}=i-1}$.
We have
\begin{align}
\notag G_{i,1}&=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{i-1}}]\indiacc{\tau_{\cC}^{i-1}=i-1}=\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{i-1}]\indiacc{\tau_{\cC}^{i-1}=i-1}\\
\label{eq:Gi1} &=g_{i-1}(\chunk[0]{X}[i-1])\indiacc{\tau_{\cC}^{i-1}=i-1}=g_{i-1,\pi}(\chunk[0]{X}[i-1])\indiacc{\tau_{\cC}^{i-1}=i-1}+R_{i,1}\eqsp.
\end{align}
Moreover, as $\tau_{\cC}^i\geqslant i$, by \eqref{eq:Gitaui=j},
\begin{align}
\notag G_{i,2}&=\sum_{j\geqslant i}\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{\tau_{\cC}^{i}}]\indiacc{\tau_{\cC}^{i-1}=i-1}\indiacc{\tau_{\cC}^i=j}\\
%\notag &=\sum_{j=i}^{n-2}\Ebb_x[f(\chunk[0]{X}[n-1])|\cF_{j}]\indiacc{\tau_{\cC}^{i-1}=i-1}\indiacc{\tau_{\cC}^i=j}+f(\chunk[0]{X}[n-1])\indiacc{\tau_{\cC}^{i-1}=i-1}\indiacc{\tau_{\cC}^i\geqslant n-1}\\
\label{eq:Gi2} &=\sum_{j=i}^{n-2}g_j(\chunk[0]{X}[j])\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i=j}+f(\chunk[0]{X}[n-1])\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i\geqslant n-1}\eqsp.
\end{align}
Let $R_{i,2}=\sum_{j=i}^{n-2}(g_j(\chunk[0]{X}[j])-g_{j,\pi}(\chunk[0]{X}[j]))\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i=j}$.
From \eqref{eq:Gi1} and \eqref{eq:Gi2},
\begin{align}
\label{eq:Gi1-Gi2} |G_{i,2}-G_{i,1}|=|&R_{i,2}-R_{i,1}+\sum_{j=i}^{n-2}(g_{j,\pi}(\chunk[0]{X}[j])-g_{i-1,\pi}(\chunk[0]{X}[i-1]))\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i=j}\\
\notag &+(f(\chunk[0]{X}[n-1])-g_{i-1,\pi}(\chunk[0]{X}[i-1]))\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i\geqslant n-1}|
% \\
% \leqslant 4&\indiacc{\tau_{\cC}^{i-1}=i-1}L\sum_{j=i+1}^{n-1}c_jr^{j-i}+\sum_{j=i}^{n-2}|(g_{j,\pi}(\chunk[0]{X}[j])-g_{i-1,\pi}(\chunk[0]{X}[i-1]))\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i=j}|\\
% &+|(f(\chunk[0]{X}[n-1])-g_{i-1,\pi}(\chunk[0]{X}[i-1]))\indiacc{\tau_{\cC}^{i-1}=i-1,\tau_{\cC}^i\geqslant n-1}|
\eqsp.
\end{align}
We bound separately all the terms in this decomposition.
First, as $\pi$ is invariant and $f\in \functionboundeddiff{\sX^n}{c}$, for any $j\in \{i+1,\ldots,n-1\}$ and any $\chunk[0]{x}[j]\in \sX^{j+1}$,
\[
|g_{j,\pi}(\chunk[0]{x}[j])-g_{i-1,\pi}(\chunk[0]{x}[i-1])|=\Ebb_{\pi}[f(\chunk[0]{x}[j],\chunk[j+1]{X}[n-1])-f(\chunk[0]{x}[i-1],\chunk[i]{X}[n-1])]\leqslant \sum_{k=i}^jc_k\eqsp.
\]
Hence,
\begin{equation}
\label{eq:Bound:Easyterms}
\begin{aligned} \sum_{j=i}^{n-2}|(g_{j,\pi}(\chunk[0]{X}[j])-g_{i-1,\pi}(\chunk[0]{X}[i-1]))|\indiacc{\tau_{\cC}^i=j}&\leqslant \sum_{j=i}^{n-2}\indiacc{\tau_{\cC}^i=j}\sum_{k=i}^jc_k=\indiacc{\tau_{\cC}^i\leqslant n-2}\sum_{k=i}^{\tau_{\cC}^i}c_k\eqsp,\\
|f(\chunk[0]{X}[n-1])-g_{i-1,\pi}(\chunk[0]{X}[i-1])|\indiacc{\tau_{\cC}^i\geqslant n-1}&\leqslant \indiacc{\tau_{\cC}^i\geqslant n-1}\sum_{k=i}^{n-1}c_k\eqsp.
\end{aligned}
\end{equation}
